---
title: "hsdsRclient -- R interfaces to HSDS (HDF5 object store)"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{hsdsRclient -- R interfaces to HSDS (HDF5 object store)}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::pdf_document:
    toc: yes
    number_sections: yes
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup,echo=FALSE,results="hide"}
suppressPackageStartupMessages({
suppressMessages({
library(hsdsRclient)
})
})
```

# Introduction

We are learning how to design R code
to work with HSDS, an object store interface
to HDF5.  The hsdsRclient package will likely merge
with rhdf5client soon (Q4 2017).

Please note the version of rhdf5client used here.
There have been issues with the handling of binary
transfer from HSDS that are not fully resolved.

```{r lkss}
sessionInfo()
```

## Some very basic benchmarking

There is a long-running server on XSEDE jetstream that
can be contacted from outside.  We demonstrate
sample-level sums for the [10x 1.3 million neuron
dataset](https://community.10xgenomics.com/t5/10x-Blog/Our-1-3-million-single-cell-dataset-is-ready-to-download/ba-p/276). 

```{r lkd}
hostPath = "/home/john/tenx_full.h5"
serverPort = "http://149.165.156.174:5101/"
txsrc = HSDS_source(serverPort, hostPath) # top level object
tx2 = HSDS_dataset(txsrc) # get a sliceable reference
apply(tx2[1:6, 1:27998],1,sum) # compute sample-level sums for first six
```

Let's benchmark this activity and then drill a little
deeper to the communication details.

```{r domic}
library(microbenchmark)
microbenchmark(apply(tx2[1:6, 1:27998],1,sum), times=5) 
```

There is a slot called `transfermode` in `tx2`:

```{r lksl}
tx2@transfermode
```

Change to 'binary':

```{r lklk}
tx2@transfermode = "binary"
microbenchmark(txs6 <- apply(tx2[1:6, 1:27998],1,sum), times=5) 
txs6
```

A modest improvement.

## Distributing the request over multiple ports

Let's do 500 samples with the simple call to binary interface.

```{r lklk2, cache=TRUE}
microbenchmark(s500 <- apply(tx2[1:500, 1:27998],1,sum), times=2) 
```

Some answers:
```{r lkans}
s500[1:6]
summary(s500)
```

It seems difficult to use multicore computing with the
communications required here.  So we use socket-based
job distribution on a multicore machine.

```{r dopar}
library(BiocParallel)
spar = SnowParam(2, type="SOCK")
register(spar)
```

With SnowParam, a complete environment
needs to be prepared on each slave.
We encapsulate the retrieval task in function
`encaps`.

```{r doenc}
encaps = function (specl = list(inds = 1:2, port = "5101"))
{
    stopifnot(all.equal(names(specl), c("inds", "port")))
    library(hsdsRclient)
    hostPath = "/home/john/tenx_full.h5"
    serverPort = sprintf("http://149.165.156.174:%s/", specl$port)
    txsrc = HSDS_source(serverPort, hostPath)
    tx2 = HSDS_dataset(txsrc)
    apply(tx2[specl$inds, 1:27998], 1, sum)
}
```

We will now query port 5101 with two parallel requests.
```{r chk2, cache=TRUE}
system.time(sans <- bplapply(list(
     list(inds=1:250, port="5101"),
     list(inds=251:500, port="5101")), encaps))
summary(unlist(sans))
```

Now use two different ports.

```{r chk3, cache=TRUE}
system.time(sans <- bplapply(list(
     list(inds=1:250, port="5101"),
     list(inds=251:500, port="5102")), encaps))
summary(unlist(sans))
```

Now use four different ports.

```{r chk4, cache=TRUE}
system.time(sans <- bplapply(list(
     list(inds=1:125, port="5101"),
     list(inds=126:250, port="5102"),
     list(inds=251:375, port="5103"),
     list(inds=376:500, port="5104")), encaps))
summary(unlist(sans))
```

Now use four different ports, full initial yield per port.

```{r chk5, cache=TRUE}
system.time(sans <- bplapply(list(
     list(inds=1:500, port="5101"),
     list(inds=501:1000, port="5102"),
     list(inds=1001:1500, port="5103"),
     list(inds=1501:2000, port="5104")), encaps))
summary(unlist(sans))
```

Do it sequentially.
```{r chk6, cache=TRUE}
system.time(sans <- lapply(list(
     list(inds=1:500, port="5101"),
     list(inds=501:1000, port="5102"),
     list(inds=1001:1500, port="5103"),
     list(inds=1501:2000, port="5104")), encaps))
summary(unlist(sans))
```
